Some lessons after the formalization of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-05-29 Independence_CH/proof_assists_isabelle.tex
One paragraph about Edinburgh LCF.

Let us briefly introduce Isabelle in the large landscape of proof
assistans (or ITPs for interactive theorem provers); we refer to the
excelent chapter by \citet{DBLP:series/hhl/HarrisonUW14} for a more
thorough recollection of the history of ITPs.

It is expected that an ITP assists the human user while mechanizing
some piece of mathematics; the interaction varies from system to
system, but a common interface consists on the system showing the
current goal and context. The user instructs the ITP to modify the
goal/context by means of tactics; a forward rule changes the context
with the intention of having in the new context a hypothesis closer to
the goal. A backward rule represents a deduction whose conclusion is
the goal, thus the new goal corresponds to the premises of the rule. A
proof is completed when the (current) goal is an instance of the
hypothesis (in the current context).

@@

To have any value at all, the system should be able to say if a tactic
makes sense (for instance, it must be forbidden to use the backward
tactic \textit{conjI} when the main connective is a
disjunction). Moreover one should be able to ascertain the validity of
the reaction of an ITP (be its rejection or its acceptance of some
alleged logical step). % Here I would merge the next paragraph.

>>>>

Let us briefly introduce Isabelle in the large landscape of proof
assistans (or ITPs for interactive theorem provers); we refer to the
excelent chapter by \citet{DBLP:series/hhl/HarrisonUW14} for a more
thorough reconstruction of the history of ITPs.

It is expected that an ITP assists the human user while mechanizing
some piece of mathematics; the interaction varies from system to
system, but a common interface consists on the system showing the
current goal and context. The user instructs the ITP to modify the
goal/context by means of tactics; a forward rule changes the context
with the intention of having in the new context a hypothesis closer to
the goal. A backward rule represents a deduction whose conclusion is
the goal, thus the new goal corresponds to the premises of the rule. A
proof is completed when the (current) goal is an instance of a
hypothesis (in the current context).

@@

To have any value at all, the system should be able to say if a tactic
makes sense (for instance, it must be forbidden to use the backward
tactic \textit{conjI} when the main connective is a
disjunction). Moreover one should be able to ascertain the validity of
the reaction of an ITP (be its rejection or its acceptance of some
alleged logical step). Edinburgh LCF~\cite{Gordon1979-qm} was an
influential proof assistant where that verification could be achieved
by checking a small \emph{kernel}; this is the only module that
constructs proofs in response to user scripts. The rest of the system
can interact with the kernel via an interface that exposes functions
guaranteeing the validity of the proofs.
